A graph-based data model for interactive evolutionary computation

### 2017-01-18 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Intro and ES-I Sections

\section{Introduction}
% Maybe an introduction on why and where to use these types of
% algorithms? - jj
% Done -Mario
Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
fitness evaluations are performed by persons within an interactive 
system.  Thus, the main loop of the EA needs the intervention of a
human for the quality assessment of each proposed solution.
IEAs have demonstrated their ability for effectively
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
as well as other products in many other domains \cite{ie1}. 
Several IECs have been developed as web based systems \cite{picbreeder},
who depend on volunteers who visit the system to interactively help
with the search, using both anonymous and registered users.
These systems have several inherent drawbacks arising from the very nature of 
the algorithms, namely, the human fatigue caused by the interaction, and
the boredom arising when users evaluate a large number of individuals 
many of which are not interesting or are very similar to each other.
IEAs also share some of the issues commonly found in volunteer based systems
\cite{sarmenta2001volunteer,web:BOINC} such as the volunteer\'s lack of accountability,
and the need to build trust between participants and application providers. 
There is also the difficulty of establishing the amount of time and resources
a volunteer is willing to spend on the system, and how they decide if they
participate or not \cite{JJ:2016}. In order to increase participation and 
engagement, and to tackle the issues mentioned above,  
IEAs must follow a user centered design, giving extensive attention 
to volunteer users, not only because their
explicit evaluation is essential, but also because the context of the 
interaction affects the system as a whole. Users interact not only with the graphical
interface, and each individual\'s phenotype, there is also the interaction
with other users, their previous evaluations and their own stream of activities
and past experiences. Thus users in this proposal are defined as dynamic entities 
interacting with an evolutionary computation having the fundamental purpose 
of evaluating individuals according to their preferences,
both explicitly or implicitly. In this work we propose giving the same
importance to users and their interactions as the population of 
individuals have in a traditional EA, even to the extend as using this
knowledge as an integral part of the algorithm it self. We argue also
that a graph model is a viable representation for this purpose.

@@

As a proof of concept, a case study is presented providing both 
conceptual and implementation details using the EvoSpace-Interactive
platform.

the EvoSpace-Interactive (ES-I) %citation
platform employs a collaborative technique,
where several registered users assign a quality assessment to a single
individual and then an aggregated fitness value is calculated. The fitness
assigned to each individual depends on the taste of each particular user, 
resulting in a many-to-many relationship between users and individuals. 
Many systems query this user-individual relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular individual, or the the user with more participation \cite{picbreeder}.
In order to do this, individuals about individuals %what? Maybe data
                                %about individuals? - JJ
must be stored, even
if they are no longer in the active population. In this work a graph database
is proposed to model the interaction between users and individuals, with
the following advantages:

\begin{enumerate}
  \item  The relationship between individuals is naturally expressed as a
  tree, or a graph when both ancestors and descendancy is considered.

  \item Relationship between users can also be modeled as a social network,
  with well stablished algorithms and metrics (cite).

  \item Many kinds of relationships between users and individuals,
  can be expressed using common actions found in social networks;
  for instance share, like, or add to collection. [ cite Activity Streams 2.0
  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]

  \item The same graph can be used to immplement a user model, or it can
  be extracted from a user\'s public profile.

\end{enumerate}

In this work..

@@

\section{EvoSpace-Interactive}
\label{sec:evospace-i}

\section{IEC Graph based model}
\label{sec:graph}

Users are a central part of an IEC system,  not only because their
explicit evaluation is essential, but also because the overall context of interaction
affects the system as a whole as users interact not only with the graphical
interface, and each individual\'s phenotype, there is also the interaction
with other users, their previous evaluations and their own history of activities
and past experiences. Thus users in this proposal are dynamic entities 
interacting with an evolutionary computation and
have the fundamental purpose of evaluating individuals according to their preferences,
both explicitly or implicitly. Their degree of participation is affected by 
the context of use. With the purpose of modeling the system, the contextual
factors are abstracted using a graph. This graph can then be used to 
enhance the system to capture the attention of users and possibly
increase their participation and mitigate the fatigue.

The current model is tailored to the EvoSpace-I functionality and capabilities
but it can be modified to handle other platforms. %A hint on how

@@

Currently there are some basic actions that
users can perform in the system, in order to interact with EvoSpace-I platform
users must login to the system through their Facebook account. 
Once a user has logged-in to the system, he or she can do several actions.
Example of actions are the following:

\begin{enumerate}
  \item  Users can invite their friends to participate through social networks .

  \item Relationship between users can also be modeled as a social network,
  with well stablished algorithms and metrics (cite).

  \item Many kinds of relationships between users and individuals,
  can be expressed using common actions found in social networks;
  for instance share, like, or add to collection. [ cite Activity Streams 2.0
  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]

  \item The same graph can be used to immplement a user model, or it can
  be extracted from a user\'s public profile.

\end{enumerate}

Another example is the sharing action, that can occur when %Something
% missing here - JJ

@@

Individuals are entities that form the population for the evolutionary
algorithm. In particular for  this work individuals are animated digital
paintings, every individual in the population is defined by a chromosome [].
This chromosome defines the behavior of the individual, so that it consists of a
vector of real numbers of fifteen genes, where each gen define a particular
behavior in painting. This individuals combined each other to generated new
individual in the population following the genetic operators such as selection,
crossover and mutation[]. An example of a individual in this case it is
illustrated in figure \ref{fig:individual}

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/individual.png}}
\caption{Individual representation.}
\label{fig:graph}       
\end{figure*}

>>>>

\section{Introduction}
% Maybe an introduction on why and where to use these types of
% algorithms? - jj
% Done -Mario
Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
fitness evaluations are performed by persons within an interactive 
system.  Thus, the main loop of the EA needs the intervention of a
human for the quality assessment of each proposed solution.
IEAs have demonstrated their ability for effectively
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
as well as other products in many other domains \cite{ie1}. 
Several IECs have been developed as web based systems \cite{picbreeder},
who depend on volunteers who visit the system to interactively help
with the search, using both anonymous and registered users.
These systems have several inherent drawbacks arising from the very nature of 
the algorithms, namely, the human fatigue caused by the interaction, and
the boredom arising when users evaluate a large number of individuals 
many of which are not interesting or are very similar to each other.
IEAs also share some of the issues commonly found in volunteer based systems
\cite{sarmenta2001volunteer,web:BOINC} such as the volunteer\'s lack of accountability,
and the need to build trust between participants and application providers. 
For the provider there is also the difficulty of establishing 
the amount of time and resources
a volunteer is willing to spend on the system, and how they decide if they
participate or not \cite{JJ:2016}. In order to increase participation and 
engagement, and to tackle the issues mentioned above,  
IEAs must follow a user centered design, giving extensive attention 
to volunteer users, not only because their
explicit evaluation is essential, but also because the context of the 
interaction affects the system as a whole. Users interact not only with the graphical
interface, and each individual\'s phenotype, there is also the interaction
with other users, their previous evaluations and their own stream of activities
and past experiences. Thus users in this proposal are defined as dynamic entities 
interacting with an evolutionary computation having the fundamental purpose 
of evaluating individuals according to their preferences,
both explicitly or implicitly. In this work we propose giving the same
importance to users and their interactions as the population of 
individuals have in a traditional EA, even to the extend as using this
knowledge as an integral part of the algorithm it self. We argue also
that a graph model is a viable representation for this purpose,  
presenting a natural way of maping the relationships found in the system. 
Most IEAs can have the following many-to-many relations:

\begin{itemize}
  \item { \tt Individual-Individual} The relationship between individuals 
  is naturally expressed as a tree, or a graph when both ancestors 
  and descendants is considered. Other semantic relations can also be considered
  for instance: similarity or composition.

  \item {\tt User-Individual}
  Many kinds of relationships between users and individuals,
  can be expressed using common actions found in social networks;
  for instance share, like, rate or add to collection [cite Activity Streams 2.0
  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]. In IEAs these
  relationships are used to assigned a single or a collection of
  fitness values to individuals [cite ES-I].

  \item {\tt User-User}
  Relationship between users can also be modeled 
  as a social network, with well stablished algorithms and metrics (cite).
  Identifyng leaders of opinion and the similarity between user\'s 
  preferences is of special importance for IEAs, as they could be used
  to recommend to users only those individuals they will probably like more.

\end{itemize}

These relations are naturaly maped to graphs, while other actions and 
properties can also be modeled as a semantic network using
basic subject–-predicate-–object expressions [cite Semantic triples].
Another reason for using a graph model is the tools available for
implementing and query the model, and the increased performance of
native graph database systems [cite Christian].

@@

A case study is presented as a proof of concept, providing both 
conceptual and implementation details of a graph data model for
the implementation of an IEA in the EvoSpace-Interactive platform. 
The graph database is then used to measure the degree of relationship
between users and their participation.

@@

\section{EvoSpace-Interactive}
\label{sec:evospace-i} 
As a case study, a IEA was implented using the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}. %citation
A brief description of the aplication is presented next, focusing
on the data elements that where ported to the graph model.

\subsection{Fitness Assignment}
\label{sec:col}
The ES-I platform employs a collaborative technique,
where several registered users assign a quality assessment to a single
individual and then an aggregated fitness value is calculated. The fitness
assigned to each individual depends on the taste of each particular user, 
resulting in a many-to-many relationship between users and individuals. 
Many systems query this user-individual relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular individual, or the the user with more participation \cite{picbreeder}.
In order to do this, data about individuals %what? Maybe data
                                %about individuals? - JJ
must be permanently stored, even
if they are no longer in the active population.

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those individuals they like, or by taking individuals
from their friend\'s collections by using the web interface depicted 
in Figure \ref{fig:web}

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=12cm,height=10cm,keepaspectratio]{img/UI_ed01.png}}
\caption{User interface ED01.}
\label{fig:web}       
\end{figure*}

At the top left corner list of Facebook friends is presented, 
to encourage users to interact with the system. In the central 
\emph{ Wall } area, an individual sampled from the population 
is shown to the user.
Here, the user can interact with the system in two ways.
He can assign a quality assessment to the individual using
a five star rating measure.
Additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special directory to store individuals a user prefers and wishes
to save. After the user finishes interacting the individual
on the Wall, he can choose to retrieve a new one from population.
At the left side, is the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, a user can browse the content of each collection and from
there share images through the social network.
When a user browses over an individual a detail pane shows how many users have
liked the individual. The pane also includes a link to the individual's 
details, the parents, genetic operators that created it, and genealogy information.

\section{IEC Graph based model}
\label{sec:graph}
The fundamental goal of this research is to develop a user-centered framework
for interactive evolutionary computation (IEC) in order to increase users’
participation and also to minimize the amount of evaluations needed for the
evolutionary process in given Web-based IEC application.

In this chapter an explanation of the proposed framework is described. The
different techniques used, such as user modeling, fuzzy logic, and human-
interaction.

This framework is presented in figure \ref{fig:uc_framework}. Each of the
entities of this framework will be explained in detail in following sections.

The current model is tailored to the EvoSpace-I functionality and capabilities
but it can be modified to handle other platforms, just by adding new clases 
of nodes.

\begin{figure*}
  \captionsetup{justification=centering,margin=2cm}
  \centering
  \setlength\fboxsep{0pt}
  \setlength\fboxrule{0.7pt}
  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/framework.png}}
  \caption{User-Centerd Framework.}
  \label{fig:uc_framework}       
\end{figure*}

%A hint on how
% done -Mario

@@

@@

Individuals are entities that form the population for the evolutionary
algorithm. In particular for  this work individuals are animated digital
paintings, every individual in the population is defined by a chromosome [].
This chromosome defines the behavior of the individual, so that it consists of a
vector of real numbers of fifteen genes, where each gen define a particular
behavior in painting. This individuals combined each other to generated new
individual in the population following the genetic operators such as selection,
crossover and mutation[].
